Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-21 PAPER.tex
Update the debate

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $P_{\rm out}$, can be thought of as a transfer function 
which, for a sampling of $P_{\rm in}$ and $P_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 
distribution. By viewing data through this signal loss lens, we may then ask the question ``what input power spectrum 
distribution could this (signal-loss affected) data come from?"

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of $e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in the main text.}

>>>>

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
\dcj{ Assuming that expansion to "leading order in $\eta$" means $\eta<<1$ then the approximation is that the 
covariance is nearly perfectly calculated from the data with only small effects due to finite samples. Thus 
correlations don't make it into the expression and foregrounds remain uncorrelated with eor or injected signals.  
I added words explaining this but it still looks strange for us to be saying that loss is linear in one section and 
then explaining in the next section exactly how cross terms were what messed us up in the first go round.}

Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $P_{\rm out}$, can be thought of as a transfer function 
which, for a sampling of $P_{\rm in}$ and $P_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 
distribution. By viewing data through this signal loss lens, we may then ask the question ``what input power spectrum 
distribution could this (signal-loss affected) data come from?"

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of 
$e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that 
couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is 
some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. 
Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other 
EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in 
the main text.}\dcj{I will admit to being a little lost here.  Doesn't changing the basis just convert the unknown f 
dot s term into other unknown terms. Isn't the whole point of Eq 24 supposed to be that we can subtract off the 
$g(x)$ to estimate $\hat{\ell}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
\ell_{\rm eor}$ not $(\ell_{\rm eor} + \rho\ell_{\rm fe})$? }
